compare the different versions of ebnm methods on missing data

### 2018-01-04 analysis/compare_ash_pn_miss.Rmd
Files commited by wflow_commit.

---
title: "apply flashr2 with OCV"
author: "wei wang"
date: 09-30
output: 
  html_document:
    code_folding: hide
---

@@

## preparation

1. creat a folder

```
mkdir Breastcancer
```

2. put your data into this folder

here we use the file 'example.mat'. the reason we use `.mat` file is because there is a matlab package in our comparision.

3. creat the wrapper function in R and matlab

### wrapper function for OCV

@@

call_method = function(Y_data,method,Kmax = 50){
  if(method == "flash"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ash_param=list(method = "fdr"),var_type = "constant")
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "PMD"){
    out = PMA::PMD(Y_data, K = Kmax)
    Y_hat =  out$u %*% diag(out$d) %*% t(out$v)
  }else if(method == "softImpute"){
    gsoft = softImpute(Y_data, rank.max = Kmax)
    Y_hat =  gsoft$u %*% diag(gsoft$d) %*% t(gsoft$v)
  }else{
    stop("the method is not implemented yet, please check it out")
  }
  return(Y_hat)
}

NSF_OCV = function(OCVindex){
  if(file.exists("NBSFAout")){
    unlink("NBSFAout", recursive= T)
  }
  system("mkdir NBSFAout")
  # write the missing index

  writeMat("./NBSFAout/OCVindex.mat", OCVindex = OCVindex)

  system("matlab -nosplash -nodesktop -r \"addpath(\'~/HG/flash/data/OCVmissflashr2/Breastcancer\'); run(\'missvalue.m\');exit;\" ")
  g_nsfa = readMat("./NBSFAout/NSFAresult.mat")
  nsfa_rmse = as.vector(g_nsfa$RMSE)
  return(nsfa_rmse)
}

@@

Here you need change the `\"addpath(\'~/HG/flash/data/OCVmissflashr2/Breastcancer\');` to your folder which contains `missvalue.m` file. this file name is `wrapper.R`

### matlab code

```{}
addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/');
addpath('~/HG/flash/data/missingvalue/methods/NBSF/nsfa-master/utils/');

% you need to change the path and the data file (not centered)
load ~/HG/flash/data/OCVmissflashr2/Breastcancer/example.mat;
Ycentered=Y-repmat(mean(Y,2),1,size(Y,2));
settings=defaultsettings();
[settings.D,settings.N]=size(Ycentered);
settings.iterations=100;
% now the missing is decide by the this is the missing index.(or you can upload the missing matrix)

load(fullfile(pwd,'NBSFAout/OCVindex.mat'))

missindex = ones(settings.D,settings.N);
k_fold = 5;
foldindex = zeros(k_fold,k_fold,2);
for k = 1:2
    for j = 1:k_fold
        for i = 1:k_fold
            foldindex(i,j,k) = OCVindex.foldindex(i + k_fold*(j - 1) + k_fold * k_fold * (k - 1));
        end
    end
end

SSE = zeros(1,k_fold);
for i = 1:k_fold
    mvmask = missindex;
    for j = 1:k_fold
        mvmask(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2))) = 0;
    end
    initialsample=init_nsfa(settings);
    [finalsample,resultstable]=nsfa(Ycentered,mvmask,initialsample,settings);
    Y_hat = (finalsample.G *  finalsample.X);
    for j = 1:k_fold
        SSE(i) = SSE(i) + ...
            sum(sum((Ycentered(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))- ...
            Y_hat(OCVindex.rowindex(:,foldindex(j,i,1)),OCVindex.colindex(:,foldindex(j,i,2)))).^2));
    end

end

RMSE = sqrt(sum(SSE)/(settings.D*settings.N));
RMSE;

save_path = fullfile(pwd,'NBSFAout/NSFAresult');
save(save_path,'RMSE'); 
```

you need change the path as well. this file name is 'missvalue.m'

### creat the run.R file to run the code

In this case, we have problem in running `flashr2` on RCC. so we just try other methods first.

```{}
library(R.matlab)
## run the code
Y_centered = readMat("~/HG/flash/data/OCVmissflashr2/Breastcancer/example.mat")
Y = Y_centered$Y
# in the matlab package of NSF, the use the centered data by rows
N = dim(Y)[1]
P = dim(Y)[2]
Y = Y - rowMeans(Y) %*% t(rep(1,P))
OCVindex = OCV_index(Y,k_fold = 5)
OCVdata = OCV_data(Y,OCVindex,k_fold = 5)
soft_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "softImpute",Kmax = 50)
# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
pmd_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "PMD",Kmax = 50)
nsfa_rmse = NSF_OCV(OCVindex)
flashG_rmse = mean(c( nsfa_rmse,pmd_rmse,soft_rmse))

result = c(flashG_rmse, nsfa_rmse,pmd_rmse,soft_rmse)
saveRDS(result, "./NBSFAout/output.rds")

this file name is `runOCV.R`

### creat folder to track the result

```
mkdir outlog
```

### creat .sbatch file

```
#!/bin/bash

#SBATCH --job-name=arrayJob
#SBATCH --output=/home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.out
#SBATCH --error=/home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/outlog/arrayJob_%A_%a.err
#SBATCH --array=1-100
#SBATCH --time=06:00:00
#SBATCH --partition=broadwl
#SBATCH --ntasks=1
#SBATCH --mem-per-cpu=2000

######################
# Begin work section #
######################

# Print this sub-job's task ID
cd /home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer
mkdir test${SLURM_ARRAY_TASK_ID}
cd test${SLURM_ARRAY_TASK_ID}
Rscript --verbose /home/weidong/HG/flash/data/OCVmissflashr2/Breastcancer/runOCV.R
```

this file is name as `OCV.sbatch`

## run OCV

now in your folder, you should have:

```
[weidong@midway2-login1 Breastcancer]$ ls
OCV.sbatch  example.mat  missvalue.m  outlog  runOCV.R  wrapper.R
```

```
sbatch OCV.sbatch
```

## to get the result

```{}
T = 100
results = matrix(NA,ncol = 4, nrow = T)
for(i in 1:T){
  test_folder = paste("test", i, sep = "")
  out_folder = "NBSFAout"
  out_file = "output.rds"
  file_name = file.path(test_folder,out_folder,out_file)
  results[i,] = try(readRDS(file_name))
}

saveRDS(results,"./boxplot.rds")
```

## plot the result

```{r}
library(ggplot2)
plot_res = function(output,title = "data",legend_position = "none"){
  rmse = as.vector(output)
  N = dim(output)[1]
  methods = rep(c("flash","NBSFA","PMD","softImpute"), each = N)
  df = data.frame(RMSE = rmse, Method = methods )
  p<-ggplot(df, aes(x=Method, y=RMSE, color=Method)) +
  geom_boxplot()+
  # geom_violin()+
  ggtitle(title) +  theme_bw()+
    theme(legend.position= legend_position, legend.text=element_text(size=15), axis.text.y = element_text(size =6))
  p
}

PT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/president_box.rds")
pp = plot_res(PT_res,"Text data")
DT_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/denoiseRtumor_box.rds")
pd = plot_res(DT_res,"Tumor data")
GZ_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/gtexzscore_box.rds")
pg = plot_res(GZ_res,"GTEx data")
BC_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/BreastCancer_box.rds")
pb = plot_res(BC_res,"Breast Cancer data")
ML_res = readRDS("../data/missingvalue/OCVtemplate/RCCres/ML100K_box.rds")
ML_res[c(2,13,17,21,29,37,62,76,77,93,95,100),] = NA
ML_res = matrix(as.numeric(ML_res),ncol = 4)
pM = plot_res(ML_res,"Movie Lens data")
```

## plots

```{r}
gridExtra::grid.arrange(pp,pd,pg,pb,pM, layout_matrix = rbind(c(1,NA,2),c(NA,5,NA),c(4,NA,3)))
```

>>>>

---
title: "compare the different versions of ebnm methods on missing data"
author: "Wei"
date: 2018-01-02
output: 
  html_document:
    code_folding: hide
---

@@

@@

call_method = function(Y_data,method,Kmax = 50){
  if(method == "ebnm_pn"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ebnm_fn = ebnm_pn,verbose = F)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else if(method == "ebnm_ash"){
    data = flashr2::flash_set_data(Y_data)
    f_greedy = flashr2::flash_add_greedy(data,Kmax=Kmax,ebnm_fn = ebnm_ash,verbose = F)
    Y_hat = f_greedy$EL %*% t(f_greedy$EF)
  }else{
    stop("the method is not implemented yet, please check it out")
  }
  return(Y_hat)
}
```

## denoise tumor data example in our paper

```{r}
set.seed(99)
library(denoiseR)
data(tumors)

Y = tumors[, -ncol(tumors)]
Y = as.matrix(Y)
N = dim(Y)[1]
P = dim(Y)[2]

OCVindex = OCV_index(Y,k_fold = 5)
OCVdata = OCV_data(Y,OCVindex,k_fold = 5)
f_pn = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_pn",Kmax = 20)
# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
f_ash = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_ash",Kmax = 20)

@@

Check the RMSE for the missing value

```{r}
c(f_ash,f_pn)
```

## Breast cancer data example in our paper

```{r}
set.seed(99)
library(R.matlab)
## run the code
Y_centered = readMat("../../ash-sfa/Rcode/postmean/flash_simulation/missingvalue/Breastcancer/example.mat")
Y = Y_centered$Y
# in the matlab package of NSF, the use the centered data by rows
N = dim(Y)[1]
P = dim(Y)[2]
Y = Y - rowMeans(Y) %*% t(rep(1,P))
OCVindex = OCV_index(Y,k_fold = 5)
OCVdata = OCV_data(Y,OCVindex,k_fold = 5)
f_pn = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_pn",Kmax = 50)
# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
f_ash = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "ebnm_ash",Kmax = 50)
```

Check the RMSE for the missing value

```{r}
c(f_ash,f_pn)
```

### MovieLens data

```{}
set.seed(11)
library(Matrix)
ml100K_data = readRDS("../../ash-sfa/Rcode/postmean/flash_simulation/missingvalue/ML100K/biscale_data.rds")
MLMatrix <- sparseMatrix(i = ml100K_data[,1],
                         j = ml100K_data[,2],
                         x = ml100K_data[,3],dims = c(943,1682))
# turn this sparse matrix into matrix in r
Y = as.matrix(MLMatrix)
Y[which(Y == 0)] = NA
OCVindex = OCV_index(Y,k_fold = 10)
OCVdata = OCV_data(Y,OCVindex,k_fold = 10)
f_pn = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "ebnm_pn",Kmax = 10)
# flashG_rmse = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 5,method = "flash",Kmax = 50)
f_ash = OCV_SSE(Y,OCVindex,OCVdata,k_fold = 10,method = "ebnm_ash",Kmax = 10)
```
